---
title: "Getting Started with ggcatppuccin"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ggcatppuccin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

ggcatppuccin provides Catppuccin color palettes and themes for ggplot2, allowing you to create beautiful, cohesive visualizations with minimal effort. The package supports all four official Catppuccin flavors: Latte, Frappé, Macchiato, and Mocha.

## Installation

```{r, eval = FALSE}
# Install from GitHub
remotes::install_github("ZhenyaKosovan/ggcatppuccin")
```

## Basic Usage

```{r setup}
library(ggplot2)
library(ggcatppuccin)
```

### Your First Plot

The simplest way to use ggcatppuccin is to add `theme_catppuccin()` and a color scale to your plot:

```{r basic-plot}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point(size = 2.5, alpha = 0.8) +
  scale_color_catppuccin() +
  theme_catppuccin() +
  labs(
    title = "Fuel Efficiency by Engine Size",
    subtitle = "Using Catppuccin Mocha theme",
    x = "Displacement (L)",
    y = "Highway MPG"
  )
```

### Choosing a Flavor

You can specify which Catppuccin flavor to use:

```{r latte-plot}
ggplot(mpg, aes(class, fill = class)) +
  geom_bar() +
  scale_fill_catppuccin(flavor = "latte") +
  theme_catppuccin(flavor = "latte") +
  guides(fill = "none") +
  labs(
    title = "Vehicle Classes",
    subtitle = "Using Catppuccin Latte theme",
    x = NULL,
    y = "Count"
  )
```

### Setting a Session Default

To avoid repeating the `flavor` argument, set a session-wide default:

```{r session-default}
set_catppuccin_flavor("macchiato")

# Now all functions use macchiato by default
ggplot(mtcars, aes(wt, mpg, color = factor(cyl))) +
  geom_point(size = 3) +
  scale_color_catppuccin() +  # Uses macchiato
  theme_catppuccin() +         # Uses macchiato
  labs(
    title = "Fuel Efficiency vs Weight",
    color = "Cylinders"
  )
```

You can still override the default by specifying `flavor` explicitly:

```{r override-default}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 2) +
  scale_color_catppuccin(flavor = "frappe") +  # Override to frappe
  theme_catppuccin(flavor = "frappe") +
  labs(title = "Iris Dataset - Frappé Flavor")
```

## Fill vs Color Scales

Use `scale_color_catppuccin()` for the `color` aesthetic and `scale_fill_catppuccin()` for the `fill` aesthetic:

```{r fill-example}
set_catppuccin_flavor("mocha")

ggplot(diamonds[sample(nrow(diamonds), 1000), ], aes(cut, fill = cut)) +
  geom_bar() +
  scale_fill_catppuccin() +
  theme_catppuccin() +
  guides(fill = "none") +
  labs(
    title = "Diamond Cut Distribution",
    x = "Cut Quality",
    y = "Count"
  )
```

## Continuous Scales

For continuous data, use the `_c` variants:

```{r continuous-scale}
ggplot(faithfuld, aes(waiting, eruptions, fill = density)) +
  geom_tile() +
  scale_fill_catppuccin_c() +
  theme_catppuccin() +
  labs(
    title = "Old Faithful Eruption Density",
    x = "Waiting Time (min)",
    y = "Eruption Time (min)"
  )
```

## Helper Geoms

The package includes convenience geoms with built-in Catppuccin colors:

```{r helper-geoms}
set_catppuccin_flavor("macchiato")

# Points default to blue
ggplot(mtcars, aes(wt, mpg)) +
  geom_catppuccin_point(size = 3) +
  theme_catppuccin() +
  labs(title = "Using geom_catppuccin_point()")

# Bars default to mauve
ggplot(mpg, aes(class)) +
  geom_catppuccin_bar() +
  theme_catppuccin() +
  labs(title = "Using geom_catppuccin_bar()")
```

## Next Steps

- Learn about [customization options](customization.html)
- See the [gallery](gallery.html) for more examples
- Learn about [R Markdown integration](rmarkdown-integration.html)
