---
title: "R Markdown and Quarto Integration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R Markdown and Quarto Integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

ggcatppuccin can style not just your plots, but your entire R Markdown or Quarto document using CSS generated from the Catppuccin palettes.

## Generating CSS

Use `catppuccin_css()` to generate a CSS string, or `write_catppuccin_css()` to write it to a file:

```{r}
library(ggcatppuccin)

# Generate CSS string
css_text <- catppuccin_css("mocha")
cat(substr(css_text, 1, 200), "...")
```

Write to a file:

```{r, eval = FALSE}
write_catppuccin_css("catppuccin.css", flavor = "mocha")
```

## Using in R Markdown

In your R Markdown YAML header:

```yaml
---
title: "My Document"
output:
  html_document:
    css: catppuccin.css
---
```

Or with a custom knit hook:

````markdown
---
title: "My Document"
output: html_document
---

```{r setup, include=FALSE}`r ''`
library(ggcatppuccin)
write_catppuccin_css("catppuccin.css", flavor = "mocha")
```
````

Then reference it:

```yaml
---
title: "My Document"
output:
  html_document:
    css: catppuccin.css
---
```

## Using in Quarto

For Quarto documents:

```yaml
---
title: "My Document"
format:
  html:
    css: catppuccin.css
---
```

Or generate dynamically:

````markdown
```{r}`r ''`
#| include: false
library(ggcatppuccin)
write_catppuccin_css("catppuccin.css", flavor = "latte")
```
````

## Coordinated Documents

Set the flavor once and use it throughout:

````markdown
```{r setup, include=FALSE}`r ''`
library(ggplot2)
library(ggcatppuccin)

# Set flavor for session
set_catppuccin_flavor("macchiato")

# Generate matching CSS
write_catppuccin_css("catppuccin.css")
```

Now all plots and the document styling will match:

```{r}`r ''`
ggplot(mtcars, aes(mpg, wt, color = factor(cyl))) +
  geom_point(size = 3) +
  scale_color_catppuccin() +  # Uses macchiato
  theme_catppuccin()           # Uses macchiato
```
````

## What Gets Styled

The generated CSS styles:

- **Body background** - Uses base color
- **Text color** - Uses text color
- **Links** - Uses accent color (mauve)
- **Code blocks** - Uses surface0 background
- **Blockquotes** - Uses accent border and surface0 background

## Example: Full Document

Here's a complete example:

````markdown
---
title: "Analysis Report"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    css: catppuccin.css
---

```{r setup, include=FALSE}`r ''`
library(ggplot2)
library(ggcatppuccin)

set_catppuccin_flavor("frappe")
write_catppuccin_css("catppuccin.css")

knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 4)
```

## Overview

This report uses the **Catppuccin Frappé** theme throughout.

## Analysis

```{r}`r ''`
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 2) +
  scale_color_catppuccin() +
  theme_catppuccin() +
  labs(title = "Iris Dataset Analysis")
```

> Note: All colors are coordinated using the Catppuccin palette.

## Code Example

```{r}`r ''`
summary(iris)
```
````

## Different Flavors for Different Purposes

- **Latte** - Light theme, good for printing
- **Frappé** - Balanced, works well on most screens
- **Macchiato** - Darker, easier on eyes in low light
- **Mocha** - Darkest, best for dark mode enthusiasts

## Tips

1. **Consistency**: Use `set_catppuccin_flavor()` once at the start
2. **File location**: Keep CSS file in the same directory as your .Rmd
3. **Version control**: Optionally add `*.css` to `.gitignore` if generating dynamically
4. **Testing**: Preview with different flavors to see which works best
